---
title: "Radiocarbon calibration"
output: learnr::tutorial
author: "Richard J. Telford"
date: "2020/10/1"
description: Calibration of single dates.
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Software for calibrating radiocarbon dates

### Web:

- [calib](calib.org)

- [OxCal](https://c14.arch.ox.ac.uk/oxcal.html)


### R packages

- rbacon
- Bchron
- oxcAAR (interface to OxCal)

Bchron is easiest to use. 
OxCal is very popular with archaeologists as it has models they can use.

## Calibration dates


### Exercise 

Calibrate and plot the dates 2450 +/- 30 BP and 2600 +/- 30 BP measured on terrestrial leaves.

First, load the `Bchron` package.
Function `BchronCalibrate` will calibrate individual dates. 
The most important arguments are `ages`, `ageSds`, and `calCurves`, which take the radiocarbon age, error, and the name of the calibration curve to use.
Either a single value or a vector can be used for each argument, but all three must be the same length.
The standard calibration curves are 'intcal20', 'shcal20' and 'marine20'.

```{r calib, exercise=TRUE, exercise.lines = 10}
library(Bchron)
cal <- BchronCalibrate(
  ages = , 
  ageSds = , 
  calCurves =     
)
summary(cal)
plot(cal)

```


```{r calib-hint}
cal <- BchronCalibrate(
  ages = c(2450, 2600), 
  ageSds = c(30, 30), 
  calCurves = rep("intcal20", 2))
summary(cal)
plot(cal)
```

1 Why are the calibrated dates' probability density functions so irregular?  
2 Why do the two dates give such different shaped calibrated dates?  
3 Calibrate the dates as if they are on marine fossils and compare the results.