---
title: "Chronologies Practical"
author: "Richard J. Telford"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

```

# Setup

Make a new project (File > New Project). This helps keep track of files and working directories.

Load the tidyverse package with `library()`

# Calibrating individual dates

Fitting an age-depth model for the Trettetjørn data.

1) Download the dates_trettetjorn.csv from mitt uib.

2) Put the file dates_trettetjorn.csv into the directory “chronology_practical/dates_trettetjorn” inside the directory where your project file is. This is because Bacon and clam are very inflexible about where the data should be.

3)  , then import the dates into R with `read_csv2()` and assign to the object `trettetjorn` with the arrow `<-`

4)    Calibrate dates individually with `IntCal::calibrate()`. To calibrate the first date, you will need code like


```{r}
library(IntCal)
calibrate(age = trettetjorn$"14C_age"[1], error = trettetjorn$error[1])
```

# Fitting a model with clam

**skip this section** - the data need reformatting a lot for it to work (unfortunately Bacon and clam need the data formatting differently).

If you still want to do it, edit the data according to instructions in the help file `?clam::clam`.

5) Fit a model with clam

```{r}
#skip this section
clam("dates_trettetjorn", coredir = "chronology_practical/")
```
Try different types of model

# Bchron

Bchron models are easier to fit than Bacon models because you can give it the data you have imported, whereas Bacon needs to find the data in a particular directory.


Load the `Bchron` package with `library()`

Fit the Bchron model with the function `Bchronology()`

Important arguments you need to set are `ages`, `ageSds`, and `positions` (i.e depth). See the examples in the help file for how to do this with `?Bchronology()`. The `calCurves` argument will default to "intcal20", which is what we want, so it does not need changing. 

After fitting the Bchron model, and assigning it to an object, you can plot it with `plot()` or make predictions for the depths you have proxy data at with `predict()`


# Running Bacon test examples.

An alternative package for generating age-depth models is `rbacon`. 

1) Load the `rbacon` package with `library(rbacon)`.
2) Run the command `Bacon()` to make an age-depth model for the ^14^C dates on the default core MSB2K. Answer "y" to the questions.
3) This will take a a few minutes (depending on how fast your computer is). When it is finished, run `agedepth()` to create a plot. 

The first plot indicates how well the Monte Carlo procedure has converged – you want this to resemble a hairy caterpillar (white noise). 

The second and third plots show the prior (green) and posterior (grey) for the sedimentation rate and memory respectively.

The main panel shows the calibrated ^14^C dates (blue) and the age depth model, where darker greys indicate more likely ages. A red line picks out the most likely age for each depth. 

4) Run `proxy.ghost()` to see the proxies with chronological uncertainty.


# Using Bacon

`dates_trettetjorn.csv` is formatted correctly for `rbacon`

The file must be saved in the expected directory or Bacon will not find it (see above for instructions).

Fit a model with Bacon. Answer “y” at the prompt.

You need to tell Bacon the name of the file (without the .csv), which must also be the name of the directory containing the file.

The `sep` argument is needed because the columns are separated by semicolons rather than the expected commas.

```{r}
Bacon("dates_trettetjorn", coredir = "chronology_practical/", sep = ";")

```


Examine the age-depth plot and check for convergence.

Bacon will put some output files into the data directory.