---
title: "Radiocarbon Dating"
author: "Richard J. Telford"
date: "2021/10/03"
output: html_document
description: Simulation of radiocarbon decay & effect of contamination on dates
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
```

## Radiocarbon decay simulation

This exercise uses a simulation of ^14^C decay.
It may take a few seconds to load.
The plot on the left shows a collection of ^14^C atoms that decay over time, the plot on the right shows the proportion of the initial activity remaining.

### Questions

- How long does it take for the ^14^C activity to fall to half the initial value?

- How long does it take for the ^14^C activity to fall from half to one quarter of the initial value?

- Why is the maximum age that ^14^C useful for about 55 000 years? 
- How many half-lives is 55 000 years?

- A fossil leaf as an ^14^C activity that is 38% of modern. 
Approximately how old is it?

```{r c14-decay-app, message = FALSE}
shinyAppDir(appDir = "14C_decay",  options = list(width = "100%", height = 700))
```

## Contamination

- What is the apparent age of an early Quaternary fossil with 1% modern contamination (where modern is 1950 i.e. 0 yr BP)?

- What is the apparent age of a modern lake sediment with 1% in-washed Carboniferous coal.

```{r contamination-app, message=FALSE}
shinyAppDir(appDir = "14C-contamination",  options = list(width = "100%", height = 700))
```

